Childcare while their mom is working

### 2022-04-03 outputs/paper.Rmd
Completed except for the title

---
title: "Childcare while their mom is working"
author: 
  - Pascal Lee Slew
  - Yunkyung Park
thanks: "Code and data are available at: https://github.com/Pascal-304/dhs_analysis."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Working mothers are often faced with the conflict of responsibility towards their children and the role they play at work. The Ghana Demographic and Health Survey (GDHS) forms part of a global initiative to collect data on family, fertility, and child health. In this paper, we used the 1998 GDHS and focused on the child care and working mothers aspect. Demographics and other aspects such as education tend to predict the child care preference. We found that families living in Greater Accra are more likely to send their child to school and the majority of working mothers look after their child by themselves. The analysis of the child care and working mothers data provides an insight on current trends and has significant implications for public policy and program decisions."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

```{r library, echo=FALSE, message=FALSE, warning = FALSE}
# install.packages('bookdown')
# install.packages("stringr")
# install.packages("ggforce")
library(stringr) 
library(dplyr)
library(tidyverse) 
library(janitor)
library(ggplot2)
library(knitr)
library(kableExtra)
library(bookdown)
library(ggforce)
```

@@

Child care matter to the development of a child. Previously, women were thought to fit the traditional role of a mother, being at home and looking after her children. This thought is being challenged with a growing presence of women in the workforce, contributing to economic activity. Women are often faced with the conflicting role they have at home and at work. However, sometimes, mothers have to work out of necessity to contribute to the financial situation of their families. Families have to find ways to raise their children. The quality of child care given tend to have positive effects on the development of the child [@care]. Having an idea of the current child care practices is important since it gives crucial information on the weaknesses of current policies and would give an insight on how to improve the situation around child care.

In this paper, we examined child care decisions of working mothers across regions, the mothers' highest education level attained, their occupation, work and employment status in Ghana. We conducted a descriptive analysis of child status and childcare preferences. We first looked at the number of working mothers from the sample population, then examined child care practices based on employment status, education level among other factors. The differences in child care practices were noticeable in different regions. We found that most working mothers tend to look after their young child by themselves and those living in Greater Accra have a higher tendency to put their child in child care centres compared to other regions. Another observation was the highest level of education attained may have an influence on the number of children they have. The proportion of working mothers with at least secondary education with no child under six years old is lower than those with at least one or more child under six.

The rest of the paper is divided into 4 sections. Section \@ref(data) explains the data source, the survey methodology and some of its key features, strengths and weaknesses. Section \@ref(results) relays the findings through graphs. Section @\ref(discussions) discusses in more detail what was found and some limitations of what we did and Section \@ref(appendix) contains a datasheet for the dataset we used.

@@

The 1998 Ghana Demographic and Health Survey was a national-level survey carried out in Ghana. This was a part of the worldwide Demographic and Health Surveys (DHS) programme and was conducted by the Ghana Statistical Service in collaboration with the Macro International Inc. It was designed to collect data on fertility, family planning, and maternal and child health. The 1998 GDHS was conducted from mid-November 1998 to mid-February 1999.

The survey frame was created using the list of Enumeration Areas (EAs) with population and household information from the 1984 Population Census. The 1998 GDHS used a two-stage stratified sampling method. At the first stage of sampling, 400 EAs were selected using systematic sampling with probability proportional to size where the selected EAs comprised 138 in the urban areas and 262 in the rural areas. A complete household listing operation was then carried out in all the selected EAs to provide a sampling frame for the second stage selection of households. At the second stage of sampling, a systematic sample of 20 households per EA was selected in the Northern, Upper West and Upper East Regions, and 15 households per EA was selected in the other regions.

@@

The GDHS was composed of 3 parts: the Household's questionnaire, the Women's questionnaire and the Men's questionnaire. All the questionnaires were modified to account for the situation in Ghana.

The survey was conducted in person. The interview staff were first trained for a period of three weeks. It was conducted by a total of 14 teams, comprising a team supervisor, a field editor, three interviewers and a driver. The response rate was 97% for individual men and women selected. Non-response was mainly due to inability to find the selected interviewee at home. Repeated callbacks were made to try to reach the selected respondents.

@@

The questionnaires were made available into the five major local languages (Akan, Ga, Ewe, Hausa, and Dagbani), making the survey accessible to a larger population. After the household listing, sport checks were conducted to make sure the work has been done correctly. Before the actual survey, a test was conducted and adjustments were made to the questionnaires based on the feedback obtained from the test. The questionnaire questions were comprehensive and detailed

@@

Some key features include background and a percent distribution of currently employed women by whether they have a child under six years of age at home according to selected background characteristics, Ghana 1998. If the respondent answered to have one or more child under six, then they were further asked to indicate the caretaker of the child such as the respondent themselves, their partner, or whether they hired a person for it, etc.

@@

It should be noted that 1.9% has not worked since the last birth, and this indicates the percentage of the respondents who were not working at the time of the survey but had been employed at some time during the last 12 months.

@@

```{r map, message=FALSE, echo = FALSE, warning=FALSE, fig.cap="Ghana administrative regions"}
# install.packages('rnaturalearth')
# install.packages('sf')
library(rnaturalearth) # we need to cite these packages as well
library(sf)
# devtools::install_github("ropensci/rnaturalearthhires")
# The st_centroid function takes the sf object and returns the center of 
# the polygon, again as a sf object.
map_ghana <- ne_states(country = 'Ghana', 
                       returnclass = 'sf')  # grab provinces within Ghana

@@

Figure \@ref(fig:map) shows a map of Ghana and its administrative regions. Ghana has 10 administrative regions, namely, Upper West, Upper East, Northern, Brong Ahato, Western, Ashanti, Central, Eastern, Volta and Greater Accra. In 1998, about 57.6% of the Ghanese population were living in rural areas [@population].

@@

Figure @\ref(fig:region) shows the number of employed mothers in the different administrative regions. There were more working mothers in Greater Accra and Ashanti while Upper West had the least number of working mothers. The reason might be because of the number of people living in the different administrative regions are different, therefore, the population sample differed from one region to another.

```{r childstatus, echo = FALSE, fig.cap = "Comparing the percentage of mothers having no child under six to one or more in both Urban and Rural areas"}
# make dataset in long format
resid_long <- residence %>% 
  select(-c(character)) %>% 
  pivot_longer(!background, names_to = "variables", values_to = "percentage") %>% 
  mutate(variables = recode(variables,
                            'no_child_under_six' = "No child under six", 
                            'one_more_child_under_six' = "One or more child under six",
                            'caretaker_respondent' = 'respondent',
                             'caretaker_partner' = 'partner',
                             'caretaker_other_relative' = 'other relative',
                             'caretaker_other_male_child' = 'other male child',
                             'caretaker_other_female_child' = 'other female child',
                             'caretaker_other' = 'other',
                             'caretaker_not_worked since_birth' = 'not worked since birth',
                             'caretaker_neighbor' = 'neighbour',
                             'caretaker_missing' = 'missing',
                             'caretaker_hired' = 'hired',
                             'caretaker_child_in_school' = 'child in school'))

@@

Figure \@ref(fig:childstatus) shows the child status of families in Urban and Rural areas. In rural areas, there is a larger percentage of working mothers having one or more child under six years old. The converse is true in Urban areas. Working mothers tend to have no child under six years old.

```{r areachildcare, echo=FALSE, fig.cap = "Comparing child caretaker preferences in Urban and Rural areas"}
# caretaker aspect
# make dataset into long format
# rename 
caretaker <- residence %>% 
  select(-c(character, no_child_under_six, one_more_child_under_six, total, number_of_employed_women, caretaker_missing)) %>% 
  pivot_longer(!background, names_to = "caretakers", values_to = "percentage") %>% 
  mutate(caretakers = recode(caretakers,
                             'caretaker_respondent' = 'respondent',
                             'caretaker_partner' = 'partner',
                             'caretaker_other_relative' = 'other relative',
                             'caretaker_other_male_child' = 'other male child',
                             'caretaker_other_female_child' = 'other female child',
                             'caretaker_other' = 'other',
                             'caretaker_not_worked_since_birth' = 'not worked since birth',
                             'caretaker_neighbor' = 'neighbour',
                             'caretaker_missing' = 'missing',
                             'caretaker_hired' = 'hired',
                             'caretaker_child_in_school' = 'child in school'))

@@

Figure @\ref(fig:areachildcare) shows the Ghanese families'caretaking preferences in both urban and rural areas. The preferred caretaking method in both urban and rural areas is the working mothers looking after their children by themselves. The next preferred method for in both areas would be to let the children be looked after by relatives. Other notable observations would be that families living in urban areas tend to put their child in school and in rural areas, families let the older female child to care for the under six years old child much more than in urban areas.

## Also comparison among the different administrative regions

```{r caretakingregions, echo = FALSE, fig.cap = "Caretaking behaviour in the different administrative regions", fig.width=7, fig.height=7}
# make dataset into long format
# recode names
child <- region %>% 
  select(-c(character, total, caretaker_missing)) %>% 
  pivot_longer(!background, names_to = "x", values_to = "percentage") %>% 
  mutate(x = recode(x,
                    'caretaker_respondent' = 'respondent',
                    'caretaker_partner' = 'partner',
                    'caretaker_other_relative' = 'other relative',
                    'caretaker_other_male_child' = 'other male child',
                    'caretaker_other_female_child' = 'other female child',
                    'caretaker_other' = 'other',
                    'caretaker_not_worked_since_birth' = 'not worked since birth',
                    'caretaker_neighbor' = 'neighbour',
                    'caretaker_missing' = 'missing',
                    'caretaker_hired' = 'hired',
                    'caretaker_child_in_school' = 'child in school',
                    'no_child_under_six' = 'no child under six',
                    'one_more_child_under_six' = 'one or more child under six',
                    'number_of_employed_women' = 'number of employed women'))

@@

Figure \@ref(fig:caretakingregions) shows the caretaking preference of Ghanese families in the different administrative regions. Working mothers tend to look after their child themselves the most in all regions. In Greater Accra, families choose to put their child in school the most compared to other administrative regions. Another observation is that the proportions of families opting to put their child in school in Northern, Upper East and Upper West are the lowest, namely 1.3%, 2.7% and 2.0% respectively. The other popular caretaking options after by the mother herself are by a relative or by another older female child.

@@

Figure \@ref(fig:chstatus) shows the child status of Ghanese families in the different administrative regions. By child status, we mean the whether the families have children under six years old. Ghanese families tend to have one or more child under six years old more than no child under six with the exception of Greater Accra. The proportion of families with one or more child under six is slightly more pronounced in Northern, Upper West, Brong Ahafo and Upper East regions.

@@

Figure \@ref(fig:motheredu) shows that there are significantly less employed women who studied at secondary school or higher. Also, it can be observed that the working mothers who did not continue their education until the secondary school are more likely to have a child under six, whereas there are more employed women who studied beyond the secondary school and have no child under six. In addition, when we take a look at the number of employed women who did not get any education, we can see that there are a lot more women with a child under six than without, and the difference between them is notably big. However, as the level of working mother's education gets higher, the gap between having a child under six and no child under six decreases, and it gets reversed when the women completed the secondary school.

@@

# plot the bar graph
edu_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage (in %)") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = 
                                   c("No_education" = "No education",
                                     "Primary" = "Primary",
                                     "Middle" = "Middle",
                                     "Secondary+" = "Secondary+")
                                 )
  )
```

Figure \@ref(fig:mothereduchildcare) demonstrates that no matter what their education level is, it is mostly the employed women who take care of their children while working. When the level of education is lower, then it is more likely for the respondent to look after their own children under six years. Over 50% of the women with no education take care of their own children whearas less than 40% of the women with secondary or higher levels of education do. As the level of the education gets higher, then the percentage of schooling their children gets higher as well. Moreover, it is worthwhile to mention that the women who completed high school are more likely to stop working after having a birth compare to the women whose education level is lower than secondary school.

@@

# plot the bar graph
work_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage (in %)") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  ) +
  facet_wrap_paginate(vars(background),
                      ncol=2,
                      nrow=2,
                      dir = "h",
                      scales = "free",
                      labeller = labeller(background = c("For_family_member" = "For family member",
                                                         "For_someone_else" = "For someone else",
                                                         "Self-employed" = "Self-employed")
                      )
  )

@@

# plot bar graph
occup_caretaker %>%
  ggplot(aes(x=caretaker, y=percentage, fill=background)) +
  geom_bar(stat = "identity", position="dodge") + 
  labs(x='Caretaker while at work', y="Percentage (in %)", fill="Occupation") +
  theme_minimal() +
  theme(legend.position='bottom') +
  coord_flip() +
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent", 
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  )
```

Figure \@ref(fig:occupationchildcare) illustrates that the percentage of the child's caretaker is highest in the respondent followed by the relative and schooling regardless of what type of jobs they have. This means that it is mostly the case that the employed women taking care of their children under six while they are working. Meanwhile, women who work in the agricultural sector are more likely to care for a child themselves than women in non-agricultural occupations. In particular, more than a half of the employed women working in the agricultural industry are the caretaker of the child under six whereas around 45% of the women working in non-agricultural sector take care of a child themselves while working. Furthermore, we can see that it is more likely for the female siblings to take care of the kids under six than the male siblings.

@@

Figure \@ref(fig:employmentstatus) illustrates that majority of the employed women work throughout the entire year. Specifically, majority of them work full week. There are few women who work occasionally. On top of that, regardless of what their employment status is, more working mothers tend to have one or more child under six.

@@

# plot the bar graph
emply_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage (in %)") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired",
        "child_in_school" = "schooling",
        "other_child" = "siblings",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = c("All_year_full_week" = "All year, full week",
                                                "All_year_part_week" = "All year, part week",
                                                "Seasonal" = "Seasonal",
                                                "Occasional" = "Occasional")
                                 )
  )

@@

We can see from Figure \@ref(fig:employmentstatuschildcare) that the percentage of the respondent being the primary caretaker of the child under six is remarkably higher than the others in all four employment status. It exceeds 40% in all four cases, and in particular, mothers working full time are more likely to look after their own children than the others. Next, it is the relatives who mainly take care of the child under six. It should be noted that there are more women sending their kids under six to school when they work all year compare to the ones who do not work throughout the entire year.

@@

We can tell from Table \@ref(tab:data) that 3,564 out of 4,843 women of age 15-49 who completed the survey are employed. In other word, around three in four women were employed during the 12 months before the survey. Figure \@ref(fig:residence) shows that the number of employed women is greater in rural area. It is worth mentioning that even though rural areas typically have low population density, rural women are somewhat more likely than urban women to be currently working. Next, Figure \@ref(fig:region) illustrates that women residing in the Upper West Region are least likely to be employed. Together with Figure \@ref(fig:map), we can tell that the women are more likely to be employed when they live in the southern part of Ghana. The northern part of Ghana including Upper West, Upper East, and Northern Region are the bottom 3 regions in Figure \@ref(fig:region) which demonstrates the distribution of the number of employed women whereas the southern part of Ghana including Greater Accra, Ashanti, and Eastern Region are the top 3.

Additionally, it is surprising to see from Figure \@ref(fig:motheredu) that the number of employed women whose education level is secondary school or higher is remarkably lower than that of the other education levels. Meanwhile, Figure \@ref(fig:occupation) shows that women are approximately twice as likely to be employed in the non-agricultural sector than in the agricultural sector. Lastly, when women are working, then Figure \@ref(fig:workstatus) tells that they are most likely to be self-employed and Figure \@ref(fig:employmentstatus) shows that they are most likely to work full-time.

>>>>

---
title: "Childcare while their mom is working"
author: 
  - Pascal Lee Slew
  - Yunkyung Park
thanks: "Code and data are available at: https://github.com/Pascal-304/dhs_analysis."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Working mothers are often faced with the conflict of responsibility towards their children and the role they play at work. The Ghana Demographic and Health Survey (GDHS) forms a part of a global initiative to collect data on family, fertility, and child health. In this paper, we used the 1998 GDHS and focused on the child care and working mothers aspect. Geographics and other aspects such as education are used to predict child care preference. We found that families living in Greater Accra are more likely to send their children to school and the majority of working mothers look after their children by themselves. The analysis of the child care and working mothers data provides insight into current trends and has significant implications for public policy and program decisions."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

```{r library, echo=FALSE, message=FALSE}
# install.packages('bookdown')
# install.packages("stringr")
# install.packages("ggforce")
library(stringr) 
library(dplyr)
library(tidyverse) 
library(janitor)
library(ggplot2)
library(knitr)
library(kableExtra)
library(bookdown)
library(ggforce)
```

@@

Child care matters to the development of a child. Previously, women were thought to fit the traditional role of a mother, being at home and looking after her children. This thought is being challenged with a growing presence of women in the workforce, contributing to economic activity. Women are often faced with the conflicting role they have at home and at work. However, sometimes, mothers have to work out of necessity to contribute to the financial situation of their families. Families have to find ways to raise their children. The quality of child care given tends to have positive effects on the development of the child [@care]. Having an idea of the current child care practices is important since it gives crucial information on the weaknesses of current policies and would give an insight on how to improve the situation around child care.

In this paper, we examined child care decisions of working mothers across regions, the mothers' highest education level attained, their occupation, work and employment status in Ghana. We conducted a descriptive analysis of child status and childcare preferences. We first looked at the number of working mothers from the sample population, then examined child care practices based on employment status, education level among other factors. The differences in child care practices were noticeable in different regions. We found that most working mothers tend to look after their young child by themselves and those living in Greater Accra have a higher tendency to put their children in child care centres compared to other regions. Another observation was the highest level of education attained may have an influence on the number of children they have. The proportion of working mothers with at least secondary education with no child under six years old is lower than those with at least one or more children under six.

The rest of the paper is divided into 4 sections. Section \@ref(data) explains the data source, the survey methodology, and some of its key features, strengths, and weaknesses. Section \@ref(results) relays the findings through graphs. Section @\ref(discussion) discusses in more detail what was found and some limitations of what we did and Section \@ref(appendix) contains a datasheet for the dataset we used.

@@

The 1998 Ghana Demographic and Health Survey was a national-level survey carried out in Ghana. It was conducted by the Ghana Statistical Service collaborated with the Macro International Inc. It was designed to collect data on fertility, family planning, and maternal and child health.

The survey frame was created using the list of Enumeration Areas (EAs) with population and household information from the 1984 Population Census. The 1998 GDHS used a two-stage stratified sampling. At the first stage of sampling, 400 EAs were selected using systematic sampling with probability proportional to size where the selected EAs comprised 138 in the urban areas and 262 in the rural areas. A complete household listing operation was then carried out in all the selected EAs to provide a sampling frame for the second stage selection of households. At the second stage of sampling, a systematic sample of 20 households per EA was selected in the Northern, Upper West and Upper East Regions, and 15 households per EA was selected in the other regions.

@@

This was a part of the worldwide Demographic and Health Surveys (DHS) programme, and the GDHS was composed of 3 parts: the Household's questionnaire, the Woman's questionnaire and the Man's questionnaire. All the questionnaires were modified to account for the situation in Ghana.

The survey was conducted in person. The interview staff were first trained for a period of three weeks. It was conducted by a total of 14 teams, comprising a team supervisor, a field editor, three interviewers and a driver. The response rate was 97% for individual men and women selected. Non-response was mainly due to inability to find the selected interviewee at home. Repeated callbacks were made to try to reach the selected respondents. The results of the survey did not disclose any confidential information that might identify the respondents. The survey was a voluntary one but it is unclear if the purpose of the survey explained to the respondents for them to make an informed decision.

@@

The questionnaires were made available in the five major local languages (Akan, Ga, Ewe, Hausa, and Dagbani), making the survey accessible to a larger population. After the household listing, sports checks were conducted to make sure the work has been done correctly. Before the actual survey, a test was conducted and adjustments were made to the questionnaires based on the feedback obtained from the test. The questions were comprehensive and detailed to collect information on the socio-economic status of the household.

@@

Some key features include background and a percent distribution of currently employed women by whether they have a child under six years of age at home according to selected background characteristics, Ghana 1998. If the respondent answered to have one or more children under six, then they were further asked to indicate the caretaker of the child such as the respondent themselves, their partner, or whether they hired a person for it, etc.

@@

It should be noted that 1.9% have not worked since their last birth, and this indicates the percentage of the respondents who were not working at the time of the survey but had been employed at some time during the last 12 months.

@@

```{r map, message=FALSE, echo = FALSE, warning=FALSE, fig.cap="Ghana administrative regions"}
# install.packages('rnaturalearth')
# install.packages('sf')
library(rnaturalearth)
library(sf)
# devtools::install_github("ropensci/rnaturalearthhires")
# The st_centroid function takes the sf object and returns the center of 
# the polygon, again as a sf object.
map_ghana <- ne_states(country = 'Ghana', 
                       returnclass = 'sf')  # grab provinces within Ghana

@@

Figure \@ref(fig:map) shows a map of Ghana and its administrative regions. Ghana has 10 administrative regions, namely, Upper West, Upper East, Northern, Brong Ahato, Western, Ashanti, Central, Eastern, Volta, and Greater Accra. In 1998, about 57.6% of the Ghanese population were living in rural areas [@population]. Greater Accra is considered the most urbanized region of Ghana.

@@

Figure @\ref(fig:region) shows the number of employed mothers in the different administrative regions. There were more working mothers in Greater Accra and Ashanti while Upper West had the least number of working mothers. The reason might be because the number of people living in the different administrative regions is different, therefore, the sample size differed from one region to another. Greater Accra is the most urbanized region which implies more work opportunities and therefore, a larger population would be concentrated there.

```{r childstatus, echo = FALSE, fig.cap = "Comparing the percentage of mothers having no child under six to one or more in both Urban and Rural areas"}
# make dataset in long format
resid_long <- residence %>% 
  select(-c(character)) %>% 
  pivot_longer(!background, names_to = "variables", values_to = "percentage") %>% 
  mutate(variables = recode(variables,
                            'no_child_under_six' = "No child under six", 
                            'one_more_child_under_six' = "One or more child under six",
                            'caretaker_respondent' = 'respondent',
                             'caretaker_partner' = 'partner',
                             'caretaker_other_relative' = 'other relative',
                             'caretaker_other_male_child' = 'other male child',
                             'caretaker_other_female_child' = 'other female child',
                             'caretaker_other' = 'other',
                             'caretaker_not_worked since_birth' = 'not worked since birth',
                             'caretaker_neighbor' = 'neighbour',
                             'caretaker_missing' = 'missing',
                             'caretaker_hired' = 'hired help',
                             'caretaker_child_in_school' = 'schooling'))

@@

Figure \@ref(fig:childstatus) shows the child status of families in Urban and Rural areas. In rural areas, there is a larger percentage of working mothers having one or more children under six years old. The converse is true in Urban areas. Working mothers tend to have no child under six years old.

```{r areachildcare, echo=FALSE, fig.cap = "Comparing child caretaker preferences in Urban and Rural areas"}
# caretaker aspect
# make dataset into long format
# rename 
caretaker <- residence %>% 
  select(-c(character, no_child_under_six, one_more_child_under_six, total, number_of_employed_women, caretaker_missing)) %>% 
  pivot_longer(!background, names_to = "caretakers", values_to = "percentage") %>% 
  mutate(caretakers = recode(caretakers,
                             'caretaker_respondent' = 'respondent',
                             'caretaker_partner' = 'partner',
                             'caretaker_other_relative' = 'other relative',
                             'caretaker_other_male_child' = 'other male child',
                             'caretaker_other_female_child' = 'other female child',
                             'caretaker_other' = 'other',
                             'caretaker_not_worked_since_birth' = 'not worked since birth',
                             'caretaker_neighbor' = 'neighbour',
                             'caretaker_missing' = 'missing',
                             'caretaker_hired' = 'hired help',
                             'caretaker_child_in_school' = 'schooling'))

@@

Figure @\ref(fig:areachildcare) shows the Ghanese families' caretaking preferences in both urban and rural areas. The preferred caretaking method in both urban and rural areas is the working mothers looking after their children by themselves. The next preferred method in both areas would be to let the children be looked after by relatives. Other notable observations would be that families living in urban areas tend to put their children in school and in rural areas, families let the older female child care for the under six years old child much more than in urban areas.

```{r caretakingregions, echo = FALSE, fig.cap = "Caretaking behaviour in the different administrative regions", fig.width=7, fig.height=7}
# make dataset into long format
# recode names
child <- region %>% 
  select(-c(character, total, caretaker_missing)) %>% 
  pivot_longer(!background, names_to = "x", values_to = "percentage") %>% 
  mutate(x = recode(x,
                    'caretaker_respondent' = 'respondent',
                    'caretaker_partner' = 'partner',
                    'caretaker_other_relative' = 'other relative',
                    'caretaker_other_male_child' = 'other male child',
                    'caretaker_other_female_child' = 'other female child',
                    'caretaker_other' = 'other',
                    'caretaker_not_worked_since_birth' = 'not worked since birth',
                    'caretaker_neighbor' = 'neighbour',
                    'caretaker_missing' = 'missing',
                    'caretaker_hired' = 'hired help',
                    'caretaker_child_in_school' = 'schooling',
                    'no_child_under_six' = 'no child under six',
                    'one_more_child_under_six' = 'one or more child under six',
                    'number_of_employed_women' = 'number of employed women'))

@@

Figure \@ref(fig:caretakingregions) shows the caretaking preference of Ghanese families in the different administrative regions. Working mothers tend to look after their children themselves the most in all regions. In Greater Accra, families choose to put their child in school the most compared to other administrative regions. Another observation is that the proportions of families opting to put their child in school in Northern, Upper East, and Upper West are the lowest, namely 1.3%, 2.7%, and 2.0% respectively. The other popular caretaking options after the mother herself are by a relative or by another older female child.

@@

Figure \@ref(fig:chstatus) shows the child status of Ghanese families in the different administrative regions. By child status, we mean whether the families have children under six years old. Ghanese families tend to have one or more children under six years old rather than no child under six with the exception of Greater Accra. The proportion of families with one or more children under six is slightly more pronounced in Northern, Upper West, Brong Ahafo, and Upper East regions.

@@

Figure \@ref(fig:motheredu) shows that there are significantly fewer employed women who studied at secondary school or higher. Also, it can be observed that the working mothers who did not continue their education until secondary school are more likely to have a child under six, whereas there are more employed women who studied beyond secondary school and have no child under six. In addition, when we take a look at the number of employed women who did not get any education, we can see that there are a lot more women with a child under six than without, and the difference between them is notably big. However, as the level of working mothers' education gets higher, the gap between having a child under six and no child under six decreases, and it gets reversed when the women complete secondary school.

@@

# plot the bar graph
edu_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage (in %)") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = 
                                   c("No_education" = "No education",
                                     "Primary" = "Primary",
                                     "Middle" = "Middle",
                                     "Secondary+" = "Secondary+")
                                 )
  )
```

Figure \@ref(fig:mothereduchildcare) demonstrates that no matter what their education level is, it is mostly the employed women who take care of their children while working. When the level of education is lower, then it is more likely for the respondent to look after their own children under six years. Over 50% of the women with no education take care of their own children whereas less than 40% of the women with secondary or higher levels of education do. As the level of education gets higher, then the percentage of schooling their children gets higher as well. Moreover, it is worthwhile to mention that the women who completed high school are more likely to stop working after having a birth compared to the women whose education level is lower than secondary school.

@@

# plot the bar graph
work_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage (in %)") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  ) +
  facet_wrap_paginate(vars(background),
                      ncol=2,
                      nrow=2,
                      dir = "h",
                      scales = "free",
                      labeller = labeller(background = c("For_family_member" = "For family member",
                                                         "For_someone_else" = "For someone else",
                                                         "Self-employed" = "Self-employed")
                      )
  )

@@

# plot bar graph
occup_caretaker %>%
  ggplot(aes(x=caretaker, y=percentage, fill=background)) +
  geom_bar(stat = "identity", position="dodge") + 
  labs(x='Caretaker while at work', y="Percentage (in %)", fill="Occupation") +
  theme_minimal() +
  theme(legend.position='bottom') +
  coord_flip() +
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent", 
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_female_child" = "other female child",
        "other_male_child" = "other male child",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  )
```

Figure \@ref(fig:occupationchildcare) illustrates that the percentage of the child's caretaker is highest among the respondent followed by the relative and schooling regardless of what type of jobs they have. This means that it is mostly the case that the employed women take care of their children under six while they are working. Meanwhile, women who work in the agricultural sector are more likely to care for a child themselves than women in the non-agricultural occupations. In particular, more than a half of the employed women working in the agricultural industry are the caretaker of the child under six whereas around 45% of the women working in non-agricultural sector take care of a child themselves while working. Furthermore, we can see that it is more likely for the female siblings to take care of the kids under six than the male siblings.

@@

Figure \@ref(fig:employmentstatus) illustrates that the majority of the employed women work throughout the entire year. Specifically, the majority of them work full week. There are few women who work occasionally. On top of that, regardless of their employment status, more working mothers tend to have one or more children under six.

@@

# plot the bar graph
emply_caretaker %>%
  ggplot(mapping=aes(x=caretaker, y=percentage, fill=caretaker)) +
  geom_bar(stat = "identity", position="dodge") + 
  theme_minimal() +
  labs(x='Caretaker while at work', y="Percentage (in %)") +
  theme(legend.position='none') +
  coord_flip() + 
  scale_x_discrete(
    labels = 
      c("respondent" = "respondent",
        "partner" = "partner",
        "other_relative" = "relative",
        "neighbor" = "neighbor",
        "hired" = "hired help",
        "child_in_school" = "schooling",
        "other_child" = "siblings",
        "not_worked_since_birth" = "not worked since birth",
        "other" = "other",
        "missing" = "missing")
  ) +
  facet_wrap(vars(background),
             dir = "v",
             scales = "free",
             labeller = labeller(background = c("All_year_full_week" = "All year, full week",
                                                "All_year_part_week" = "All year, part week",
                                                "Seasonal" = "Seasonal",
                                                "Occasional" = "Occasional")
                                 )
  )

@@

We can see from Figure \@ref(fig:employmentstatuschildcare) that the percentage of the respondent being the primary caretaker of a child under six is remarkably higher than the others in all four employment statuses. It exceeds 40% in all four cases, and in particular, mothers working full time are more likely to look after their own children than the others. Next, it is the relatives who mainly take care of the child under six. It should be noted that there are more women sending their kids under six to school when they work all year compared to the ones who do not work throughout the entire year.

@@

Table \@ref(tab:data) illustrates that 3,564 out of 4,843 women of age 15-49 who completed the survey are employed. In other words, around three in four women were employed during the 12 months before the survey. Figure \@ref(fig:residence) shows that there are more employed women in the rural area, and this is because Ghana's population is predominantly rural. 34% of the country's population was reported to be residing in urban areas, and 66% percent were living in rural areas [@ghana1998].

Next, Figure \@ref(fig:region) illustrates that women residing in the Upper West Region are least likely to be employed. Together with Figure \@ref(fig:map), we can tell that women are more likely to be employed when they live in the southern part of Ghana. The northern part of Ghana including Upper West, Upper East, and Northern Region are the bottom 3 regions in Figure \@ref(fig:region) which demonstrates the distribution of the number of employed women whereas the southern part of Ghana including Greater Accra, Ashanti, and Eastern Region are the top 3.

Additionally, it is surprising to see from Figure \@ref(fig:motheredu) that the number of employed women whose education level is a secondary school or higher is remarkably lower than that of the other education levels. Meanwhile, Figure \@ref(fig:occupation) shows that women are approximately twice as likely to be employed in the non-agricultural sector than in the agricultural sector. Lastly, when women are working, then Figure \@ref(fig:workstatus) tells that they are most likely to be self-employed and Figure \@ref(fig:employmentstatus) shows that they are most likely to work full-time.
